Paper

### 2020-02-24 Paper.Rmd
improving plots

---
title: "Paper"
author: "Sarah Killcoyne"
date: "2019-11-13"
output: 
  html_document:
    fig_height: 5
    fig_width: 5
    toc: yes
    toc_depth: 4
editor_options: 
  chunk_output_type: console
---

@@

```{r p53roc, echo=F, fig.height=2, fig.width=2}
p53samples = subset(patient.info, !is.na(p53.Status), select=c('Samplename','p53.Status','Status'))
p53roc = pROC::roc(as.numeric(p53samples$Status)-1, as.integer(p53samples$p53.Status)-1)

@@

#### By pathology?
```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
p53Path = pg.samp %>% dplyr::filter(Status == 'P' & !is.na(p53.Status)) %>% 
  dplyr::group_by(Pathology,p53.Status) %>% 
  dplyr::summarise(Freq = length(Pathology)) %>% dplyr::mutate(Status = 'p53 IHC')

p53Path = full_join( pg.samp %>% dplyr::filter(Status == 'P' & !is.na(p53.Status)) %>% 
                       dplyr::group_by(Pathology) %>% 
                       dplyr::summarise(nPath = length(Pathology)), p53Path, by='Pathology') %>% 
  dplyr::mutate(ratio = round(Freq/nPath,2))

rp = plot.risk.by.path(pg.samp %>% ungroup) + labs(title='Samples classified by pathology', y='% Classified')
rp
ggsave(filename='plots/pred_by_path.png', plot=(rp + theme(legend.position = 'none')), width=7, height=6, units='in', dpi=300)
```

```{r p53IHC, echo=F, warning=F, message=F, fig.height=8, fig.width=8}
#p53Path$Risk[p53Path$Risk == 'High'] = 'Aberrant'
#p53Path$Risk[p53Path$Risk == 'Low'] = 'Normal'

#p53Path$p53.Status = factor(p53Path$p53.Status, levels=c('Aberrant', 'Normal'), ordered = F)

p53Path$p53.Status = factor(p53Path$p53.Status, ordered = T)
p53Path =arrange(p53Path, Pathology, p53.Status)
col = brewer.pal(11, "PRGn")[c(10,2)]
rp = ggplot(p53Path, aes(Pathology, ratio*100, group=Pathology, fill=p53.Status)) + geom_bar(stat='identity', position = position_stack()) +
  scale_fill_manual(values=col, name='p53 IHC', labels=c('Normal','Aberrant')) + 
  geom_text(aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101), nudge_y=1) +
  geom_text(aes(label=paste(ratio*100,'%',sep='')), position=position_stack(vjust = 0.5), size=5,color='white') +
  plot.theme + labs(title='Samples predicted by aberrant p53 IHC', y='% Predicted', x='Pathology') + 
  theme(legend.position = 'bottom', text = element_text(size=14))
rp
ggsave(filename='plots/p53_pred.png', plot=rp, width=6, height=5, units='in', dpi=300)
```

@@

    p = BarrettsProgressionRisk::patientRiskTilesPlot(patient, 'months.before.final','rev') + theme(legend.position = 'none') + labs(x='months',y='location') + theme(plot.title = element_text(face = 'bold', size=16))

    ggsave(paste0('plots/heatmaps/',unique(patient$Status),'/pt_',pt,'.png'), plot = p, height=length(unique(patient$GEJ.Distance))+1, width=length(unique(patient$months.before.final))+1, units='in', dpi=300, limitsize=F)

@@

  p = BarrettsProgressionRisk::patientRiskTilesPlot(patient, 'months.before.final', 'rev') + labs(x='months', y='location') + theme(legend.position = 'none',plot.title = element_text(face = 'bold', size=16))

    ggsave(paste0('plots/validation/heatmaps/',unique(patient$Status),'/pt_',pt,'.png'), plot = p, height=length(unique(patient$GEJ.Distance))+1, width=length(unique(patient$months.before.final))+1, units='in', dpi=300, limitsize=F) 
}

@@

>>>>

---
title: "Paper"
author: "Sarah Killcoyne"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  html_document:
    fig_height: 5
    fig_width: 5
    toc: yes
    toc_depth: 4
editor_options: 
  chunk_output_type: console
---

@@

```{r p53roc, echo=F, fig.height=2, fig.width=2}
p53samples = subset(patient.info, !is.na(p53.Status), select=c('Samplename','p53.Status','Status', 'Pathology'))
p53roc = pROC::roc(as.numeric(p53samples$Status)-1, as.integer(p53samples$p53.Status)-1)

@@

#### By pathology?
```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
rp = plot.risk.by.path(pg.samp %>% ungroup) + labs(title='Samples classified by pathology', y='% Classified')
rp
ggsave(filename='plots/pred_by_path.png', plot=(rp + theme(legend.position = 'none')), width=7, height=6, units='in', dpi=300)
```

```{r p53IHC, echo=F, warning=F, message=F, fig.height=8, fig.width=8}
p53Path = pg.samp %>% dplyr::filter(Status == 'P' & !is.na(p53.Status)) %>% 
  dplyr::group_by(Pathology,p53.Status) %>% tally(name='Freq') %>% ungroup %>% 
  group_by(Pathology) %>% dplyr::mutate(nPath = sum(Freq), Status = 'p53 IHC', ratio = round(Freq/nPath,2))

p53Path$p53.Status = factor(p53Path$p53.Status, ordered = T)
p53Path = arrange(p53Path, Pathology, p53.Status)
col = brewer.pal(11, "PRGn")[c(10,2)]
rp = ggplot(p53Path, aes(Pathology, ratio*100, group=Pathology, fill=p53.Status)) + geom_bar(stat='identity', position = position_stack()) +
  scale_fill_manual(values=col, name='p53 IHC', labels=c('Normal','Aberrant')) + 
  geom_text(aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101), nudge_y=1) +
  geom_text(aes(label=paste(ratio*100,'%',sep='')), position=position_stack(vjust = 0.5), size=5,color='white') +
  plot.theme + labs(title='p53 IHC in samples from progressive patients', y='% samples', x='Pathology') + 
  theme(legend.position = 'bottom', text = element_text(size=14))
rp
ggsave(filename='plots/p53_pred.png', plot=rp, width=6, height=5, units='in', dpi=300)
```

@@

    p = BarrettsProgressionRisk::patientRiskTilesPlot(patient, 'months.before.final','rev', 0.5) + theme(legend.position = 'none') + labs(x='months',y='location') + theme(text = element_text(size = 10), plot.title = element_text(face = 'bold', size=16), axis.text = element_text(size = 10))

      h = (length(unique(patient$GEJ.Distance))+1)*0.5
  if (length(unique(patient$GEJ.Distance)) <= 1) h = (length(unique(patient$GEJ.Distance))+1)*0.8
  w = (length(unique(patient$months.before.final))+1)*0.5
  if (length(unique(patient$months.before.final)) <= 1) w = (length(unique(patient$months.before.final))+1)*0.8

    ggsave(paste0('plots/heatmaps/',unique(patient$Status),'/pt_',pt,'.png'), plot = p, height=h, width=w, units='in', dpi=300, limitsize=F)

@@

  p = BarrettsProgressionRisk::patientRiskTilesPlot(patient, 'months.before.final', 'rev', 0.5) + labs(x='months', y='location') + theme(legend.position = 'none',plot.title = element_text(face = 'bold', size=16), text = element_text(size=12), axis.text = element_text(size=12))

    h = (length(unique(patient$GEJ.Distance))+1)*0.8
  if (length(unique(patient$GEJ.Distance)) <= 1) h = (length(unique(patient$GEJ.Distance))+1)
  w = (length(unique(patient$months.before.final))+1)*0.8
  if (length(unique(patient$months.before.final)) <= 1) w = (length(unique(patient$months.before.final))+1)

    ggsave(paste0('plots/validation/heatmaps/',unique(patient$Status),'/pt_',pt,'.png'), plot = p, height=h, width=w, units='in', dpi=300, limitsize=F) 
}

@@

# CoxHD with clinical factors
```{r, eval=F}
library(CoxHD)

time<-function(x) {
  for (i in 1:nrow(x)) {
    if (i == 1) {
      x[i,'tstart'] = 0
      x[i, 'tstop'] = x[(i+1), 'Endoscopy.Year'] - x[i, 'Endoscopy.Year']
      if (is.na(x[i,'tstop']) || x[i,'tstop'] == 0) x[i,'tstop'] = 0.5
      next
    }

    x[i,'tstart'] = x[(i-1), 'tstop']

    # Final sample is different -- will leave it out later
    if (i == nrow(x)) {
      x[i, 'tstop'] = x[i,'tstart'] + 0.5
      break
    }

    x[i, 'tstop'] =  x[i,'tstart'] + x[(i+1), 'Endoscopy.Year'] - x[i, 'Endoscopy.Year']

    # Samples taken in the same year
    if ( x[i, 'Endoscopy.Year'] == x[(i+1), 'Endoscopy.Year'] )
      x[i,'tstop'] = x[i,'tstop'] + 0.5 
  }
  return(x[,c('tstart','tstop')])
}

coxDF = do.call(bind_rows, purrr::map(unique(pg.samp$Patient), function(pt) {
  df = pg.samp %>% filter(Patient == pt) %>% arrange(Endoscopy.Year)
  df %>% bind_cols(time(df))
})) %>%  dplyr::select(Samplename, Patient, Status, Risk, Prediction, Pathology, tstart, tstop, Age.at.diagnosis, Smoking, Sex, Maximal) %>% 
  mutate(Diagnosis = as.integer(Pathology %in% c('HGD','IMC')) ) %>%  dplyr::select(Samplename, Patient, Status, tstart, tstop, Diagnosis, everything() )

coxDF = bind_cols(coxDF, as_tibble(dysplasia.df[coxDF$Samplename,features$label])) %>% 
  rename_at( vars(matches('\\d+:')), list(~paste0('chr',gsub(':|-','.', .))) ) %>% 
  mutate(Smoking = as.integer(Smoking)-1, Sex = as.integer(Sex)-1, Pathology = as.integer(Pathology)-1) %>% ungroup

coxSM = CoxHD::StandardizeMagnitude(dplyr::select(coxDF, -matches('Samplename|Patient|Status|Risk|Prediction|Pathology|tstart|tstop|Diagnosis')))
coxSM = bind_cols(dplyr::select(coxDF, Samplename, tstart, tstop, Diagnosis), as_tibble(coxSM))

surv = with(coxSM, Surv(tstart, tstop, Diagnosis))

groups = factor(c(rep('clinical',4),c(rep('genomic', nrow(features)))))
rfx = CoxRFX(as.data.frame(coxSM[,-c(1:4)]), surv, groups, verbose=T, nu=1)
summary(rfx)

plot(rfx)

wt = WaldTest(rfx)
survConcordance(surv ~ predict(rfx))

wt$label = sub('\\.', '-', sub('\\.', ':', sub('chr', '', rownames(wt))))
wt = merge(wt, features[,c('label','hazards', 'coef','stability')], by='label', all=T)

wt[is.na(wt$hazards), c('hazards', 'coef.y')] = 0
plot.waldtest(wt) + labs(x='RFX coef',y='GLM coef', title='RFX vs GLM coefficients', subtitle='Labeled points have a significant p-value in RFX (except BE/ID)')

#gcols = colnames(coxDF)
#cols = features$label

coxSM = cbind(coxDF[,c(1:3)], StandardizeMagnitude(coxDF[,-c(1:3)]))
